---
title: "Packageless coding; Programming with base R"
author: "Logan Ossentjuk"
date: "11/23/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Objects and classes

## Vectors

### Creating a vector

```{r}
# Numeric vector 

my_vector <- vector(mode = "numeric", length = 100)

# Logical vector

my_second_vector = vector(mode = "logical", length = 20)
```

### Vector properties

#### Length 


```{r}
#Length of my vector 
length(my_vector)
length(my_second_vector)

```
 #### Class 
 
```{r}
#WHat's in my vector?
class(my_vector)
class(my_second_vector)
```
 
 #### Indexing (accessing) vectors
 
```{r}
# using c to create a vector 
my_vector <- c(1:50)

#What is the first element?
my_vector[1]

#what is the last element?
my_vector[-1]

#what is the seventh element? 
my_vector[7]

#select a range from a vector
my_vector[3:7]
```

## Matrices 

### Creating matrices 

```{r}
#Creating a 10 x 10 matrix full of NA's

my_matrix <- matrix(data = 1:100, nrow = 20, ncol = 5)
```

 ### Matrix Properties 
 
 #### Dimensions 
 
```{r}
#What are the dimensions
dim(my_matrix) #first value is rows, second is columns
```
 
 #### Length
 
```{r}
#How many elements are there? 
length(my_matrix)
```
 
 #### Class
```{r}
class(my_matrix)
```
 
 #### Indexing 
 
```{r}
#index a row 
my_matrix[2,]

#index a column
my_matrix[,3]

#index a given cell 
my_matrix[8,3]

# Index and replace 

my_matrix[8,3] <- 100000

#Replace an entire column 

my_matrix[,5] <- 10
```
  
# Control Flow 

## 'for' loops

variable (counter), sequence, expression

```{r}
for (i in 1:10) {
  print("Hello world!") 
}
```

# Generating random numbers 

```{r}
#draws five observations from a normal distribution  within one standard deviation of the given mean
rnorm(n = 5, mean = 10, sd = 2)
```


# Let's combine all of the above 

We will use 'numeric' to create anempty vector 
We will use 'for' to iterate along items of that empty vector 
We will use '[]' to access each slot and fill in that empty slot 
We will use 'rnorm' to sample values from a random normal distribution 
We will plot the trend 

$$N_{t}=N_{t-1} + r\left(1- \frac{N_{t-1}}{K}\right)$$
$$
r \sim norm(\mu = 1.5, \sigma = 0.8)
$$


## Writing the code down 


```{r}
# Set up model parameter
K <- 100
nsteps <- 100
mu <- 0.4
sigma <- 1

# Set up model objects 

N <- vector(mode = 'numeric', length = nsteps)
N[1] <- 25

# Model it 
for(t in 2:nsteps) {
  r <- rnorm(n = 1, mean = mu, sd = sigma)
  N[t] <- N[t-1] + r * (1 - (N[t-1] / K))
}

# Visualize it
plot(N)
```

# Something more 

We will use 'matrix' to create an empty matrix (rows are time, columns are runs)
We will repeat the process

```{r}
# Set up model parametesr
K <- 100
nsteps <- 100
mu <- 0.4
sigma <- 1

# Set up model objects 

N <- matrix(data = NA, nrow = nsteps, ncol = 100)
N[1,] <- 25

# Model it 
for(t in 2:nsteps) {
  r <- rnorm(n = 100, mean = mu, sd = sigma)
  N[t,] <- N[t-1,] + r * (1 - (N[t-1,] / K))
}

# Visualize it
matplot(N, type = "l")
```

# Bring in your friends:

Convert it to things we know about: `ggplot`

```{r}
#Load tidyverse
library(tidyverse)

# Create data.frame
my_data <- as_tibble(N) %>% 
  mutate(time = 1:nsteps) %>% 
  gather(run, N, -time)

# Visualize it 
ggplot(data = my_data, mapping = aes(x = time, y = N)) +
  geom_line(aes(group = run), size = 0.1, alpha = 0.5) +
  stat_summary(geom = "line", fun = mean, color = "red", size = 1) + 
  theme_bw() +
  labs(x = "Time (years)", y = 'Population Size (N)')

```

